---
title: "Reading data"
author: "Ricardo Mayer"
date: "February 13, 2017"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "./../")
knitr::opts_chunk$set(echo = TRUE)
```

## Reading data

### Some criteria

1.  Import directly from online sources whenever available. This implies that in the future, when I have worked out how to make direct queries and downloads from the Cepalstat API service, I will not use donwloaded excel files from cepalstat web facility.

2.  Prefer cvs output over excel. Files are lighter to move around (e.g. up and down the github repository) and more robust to import.

3. When importing excel files, prefer the `read_xls` package: it is simple, fast and it's boud to become standard (it's the package used by RStudio IDE for importing excel files). When the region within the sheet is not so simple (i.e. specifying how many rows to skip it's not enough), then use the `XLConnect` package. It's a great package, specially to write back to excel, that let you specify the region (even named regions) you want to import. Unfortunately, it's reliance on Java makes it less easy to install, specially without admin privileges.

4. Store retrieved data in the `raw_data` directory and treat it as read-only files.

5. Make no changes to the read files, with the exception --out of convenience-- of dealing with warning messages that appear during read-time (and hence probably forgotten later) and that make no sustantive changes to data. For example, when reading cepalstat csv files, a spurious column is created and you receive a woring message that notify of a newly created X8 column, added at the end of the data set. To remove this and therefore, to preserve the original data content, in the reading script  is OK to do it inside the reading script.

5. Save the read raw data files, whatever its file format, inside the `produced_data` directory. Prefer `.rda` or `.rds` format to save imported data. 

6. Use the `read_` prefix to name any script used to read or retrieve data. Keep those 'R' scripts in the `scripts` directory.

7. Keep the reading procedure and data wragling/munging separated. At most, write necessary code to hold several sheets inside a single list, but leave any renaming, selection, filtering, joining, etc. to a separate script.  


### Preliminaries: country dictionaries
With the help of the `countrycode`  package we create custom 
dataframes to hold iso2c, iso3c, english names and spanish names for the complete list 
of latinoamerican and caribbean members of ECLAC (33 countries) and a more restricted subset of 20 economies.

The `countrycode` package let us to change easly between codes and names, which is useful to retrive information and to construct consistent ID keys to merge datasents down the road.

In this shorter list we have excluded all caribbean member countries except Dominican Republic and 
included all members from Central and South America , with the exception of Belize and Suriname.

At the end of this part, we save these dataframes to call them later in subsequent scripts.

The process is carried in the `generate_country_dicts.R` script and reproduced here:

```{r}
library(countrycode)
library(tidyverse)

ca_sa_region = c("Central America", "South America")
sa_not_cepal = c("FLK", "GUF")
cepal_caribbean = c("ATG","BHS", "BRB", "CUB", "DMA",
                    "DOM", "GRD", "HTI", "JAM", "KNA",
                    "LCA", "VCT", "TTO")

cepal_33_countries <- countrycode_data %>% 
                select(iso2c, iso3c, country.name.en, country.name.es, region) %>% 
                filter(region %in% ca_sa_region | iso3c %in% cepal_caribbean) %>% 
                filter(! iso3c  %in%  sa_not_cepal) 
                

# fix the Trinidad yTobago entry and change it to Trinidad y Tabago
cepal_33_countries$country.name.es[cepal_33_countries$iso3c == "TTO"] <- "Trinidad y Tabago"


carib_minus_dom = c("ATG", "BHS", "BRB", "DMA", "GRD", 
                    "HTI", "KNA", "LCA", "TTO", "VCT")

other_to_drop = c("GUY", "SUR", "BLZ") 

# without caribbean (except cuba and dom rep) and without belize, suriname and guyana
cepal_20_countries = cepal_33_countries %>% 
  filter(!iso3c %in% carib_minus_dom) %>% 
  filter(!iso3c %in% other_to_drop)


save(cepal_33_countries, file = "./produced_data/cepal_33_countries")
save(cepal_20_countries, file = "./produced_data/cepal_20_countries")

```



### Read cepalstat's data

What I did here was to use the advanced query utility ("combinar indicadores/interfaz avanzada"), see [here](http://interwp.cepal.org/cepalstat/engine/index.html), to download bulkier .csv files with as many indicators as possible. Basically, all the indicators inside the Economics branch and various related to employment found in the Social branch. Since this implies to consider monthly, quaterly and topical data, the long format result in a very large number of row. This is still the best way to store data, IMHO, but later will need extensive use of filtering and grouping operations to analize the datasets. 

List of file and data:

- *cepalstat_sector_real_dolares_anual_completo.csv* : System of National Account annual data in US dollars. Several emasures of GDP and GDP growth, national income, gross formation of fixed capital, external gap, GDP deflator and openness.

- *cepalstat_desempleo.csv*

- *cepalstat_empleo.csv*

- *cepalstat_remuneraciones.csv*

- *cepalstat_sector_financiero_monetario.csv*: nominal interest rates, policy rate, M1, cash, deposits, M3

- *cepalstat_sector_publico2.csv*: Tax revenue, income tax rates, VAT, Operations, public debt

- *cepalstat_BP_anual*: Balance of Payments

- *cepalstat_BP_trimestral.csv*

- *cepalstat_indicadores_derivados_de_la_BP.csv*: value of X, M, transfers, rents, assests, liabilities

- *cepalstat_exp_imp_servicios.csv*

- *cepalstat_deuda_externa.csv*

- *cepalstat_comercio_intrarregional.csv*

- *cepalstat_exp_imp_pro_ppal_part_1_of_2.csv*: principal goods

- *cepalstat_exp_imp_pro_ppal_part_2_of_2.csv*

- *cepalstat_exp_imp_totales_mensuales.csv*

- *cepalstat_exp_prim_manuf.csv*

- *cepalstat_indic_vol_precios_imp_exp.csv*

- *cepalstat_tipo_de_cambio.csv*

- *cepalstat_ipc_ipm_mensual.csv*

- *cepalstat_ipc_ipm_anual_nivel_y_var.csv*

- *cepalstat_agricultura.csv*

- *cepalstat_mineria_manufactura.csv*

- *cepalstat_precios_combustibles.csv*

- *cepalstat_turismo.csv*

- *cepalstat_sector_real_mn_na_trimestral.csv*


Here is the initial readings of the following data. I had to experiment a little to find the right encoding to read these files (it was `windows-1252`). For the first reading only, I'll display here the message issued by `read_delim`, that form the reason to apply the correction that you will find later in this document and supress them here on.

```{r read_csv_files, cache=TRUE}
library(tidyverse, quietly = TRUE)
cepalstat_sector_real_dolares_completo <- 
                    read_delim("./raw_data/cepalstat_sector_real_dolares_anual_completo.csv",
                    ";", escape_double = FALSE,
                    trim_ws = TRUE, 
                    locale = locale("es", encoding="windows-1252"))

```

That mysterious new column called `X10` is entirely composed of NAs, it seems to be an artifact of reading and I'll proceed to remove it later. This chunk reveals the content of `X10`:

```{r show_X10}
summary(as.factor(cepalstat_sector_real_dolares_completo$X10))
```

Here are the rest of the data rweading calls:

```{r read_rest, cache=TRUE, warning=FALSE, message=FALSE, tidy=FALSE}
cepalstat_desempleo <- read_delim("./raw_data/cepalstat_desempleo.csv", 
                                  ";", escape_double = FALSE, 
                                  trim_ws = TRUE,
                                  locale = locale("es", encoding="windows-1252"))

cepalstat_empleo <- read_delim("./raw_data/cepalstat_empleo.csv", 
                               ";", escape_double = FALSE, 
                               trim_ws = TRUE,
                               locale = locale("es", encoding="windows-1252"))

cepalstat_remuneraciones <- read_delim("./raw_data/cepalstat_remuneraciones.csv", 
                                       ";", escape_double = FALSE, 
                                       trim_ws = TRUE,
                                       locale = locale("es", encoding="windows-1252"))

cepalstat_sector_financiero_monetario <- read_delim(
                                      "./raw_data/cepalstat_sector_financiero_monetario.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_sector_publico <- read_delim("./raw_data/cepalstat_sector_publico2.csv", 
                                        ";", escape_double = FALSE, 
                                        trim_ws = TRUE, 
                                       locale = locale("es", encoding="windows-1252"))

cepalstat_BP_anual <- read_delim("./raw_data/cepalstat_BP_anual.csv", 
                                 ";", escape_double = FALSE, 
                                 trim_ws = TRUE, 
                                 locale = locale("es", encoding="windows-1252"))

cepalstat_BP_trimestral <- read_delim("./raw_data/cepalstat_BP_trimestral.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_indicadores_derivados_de_la_BP <- read_delim(
                                  "./raw_data/cepalstat_indicadores_derivados_de_la_BP.csv", 
                                  ";", escape_double = FALSE, 
                                  trim_ws = TRUE, 
                                  locale = locale("es", encoding="windows-1252"))

cepalstat_exp_imp_servicios <- read_delim("./raw_data/cepalstat_exp_imp_servicios.csv", 
                                          ";", escape_double = FALSE, 
                                          trim_ws = TRUE, 
                                          locale = locale("es", encoding="windows-1252"))

cepalstat_deuda_externa <- read_delim("./raw_data/cepalstat_deuda_externa.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_comercio_intrarregional <- read_delim(
                                     "./raw_data/cepalstat_comercio_intrarregional.csv", 
                                     ";", escape_double = FALSE, 
                                     trim_ws = TRUE, 
                                     locale = locale("es", encoding="windows-1252"))

cepalstat_exp_imp_grandes_cat <- read_delim("./raw_data/cepalstat_exp_imp_grandes_cat.csv", 
                                            ";", escape_double = FALSE, 
                                            trim_ws = TRUE, 
                                            locale = locale("es", encoding="windows-1252"))

cepalstat_exp_imp_pro_ppal_part_1_of_2 <- read_delim(
                                    "./raw_data/cepalstat_exp_imp_pro_ppal_part_1_of_2.csv", 
                                    ";", escape_double = FALSE, 
                                    trim_ws = TRUE, 
                                    locale = locale("es", encoding="windows-1252"))

cepalstat_exp_imp_pro_ppal_part_2_of_2 <- read_delim(
                                    "./raw_data/cepalstat_exp_imp_pro_ppal_part_2_of_2.csv", 
                                    ";", escape_double = FALSE, 
                                    trim_ws = TRUE, 
                                    locale = locale("es", encoding="windows-1252"))

cepalstat_exp_imp_totales_mensuales <- read_delim(
                                      "./raw_data/cepalstat_exp_imp_totales_mensuales.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_exp_prim_manuf <- read_delim("./raw_data/cepalstat_exp_prim_manuf.csv", 
                                       ";", escape_double = FALSE, 
                                       trim_ws = TRUE, 
                                       locale = locale("es", encoding="windows-1252"))

cepalstat_indic_vol_precios_imp_exp <- read_delim(
                                      "./raw_data/cepalstat_indic_vol_precios_imp_exp.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_tipo_de_cambio <- read_delim("./raw_data/cepalstat_tipo_de_cambio.csv", 
                                       ";", escape_double = FALSE, 
                                       trim_ws = TRUE, 
                                       locale = locale("es", encoding="windows-1252"))

cepalstat_ipc_ipm_mensual <- read_delim("./raw_data/cepalstat_ipc_ipm_mensual.csv", 
                                       ";", escape_double = FALSE, 
                                       trim_ws = TRUE, 
                                       locale = locale("es", encoding="windows-1252"))

cepalstat_ipc_ipm_anual <- read_delim("./raw_data/cepalstat_ipc_ipm_anual_nivel_y_var.csv", 
                                       ";", escape_double = FALSE, 
                                       trim_ws = TRUE, 
                                       locale = locale("es", encoding="windows-1252"))

cepalstat_agricultura <- read_delim("./raw_data/cepalstat_agricultura.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_mineria_manuf <- read_delim("./raw_data/cepalstat_mineria_manufactura.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_precios_combustibles <- read_delim("./raw_data/cepalstat_precios_combustibles.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_turismo <- read_delim("./raw_data/cepalstat_turismo.csv", 
                                      ";", escape_double = FALSE, 
                                      trim_ws = TRUE, 
                                      locale = locale("es", encoding="windows-1252"))

cepalstat_sector_real_mn_trimestral <-
  read_delim(
  "./raw_data/cepalstat_sector_real_mn_na_trimestral.csv",
  ";",
  escape_double = FALSE,
  trim_ws = TRUE,
  locale = locale("es", encoding = "windows-1252")
  )
```


Next, I'll remove the extra `X` column added to each dataset

```{r removing_X_columns, warning=FALSE, message=FALSE}
library(tidyverse, quietly = TRUE)
new_cols = ncol(cepalstat_sector_real_dolares_completo) - 1
cepalstat_sector_real_dolares_completo = select(cepalstat_sector_real_dolares_completo, 
                                                1:new_cols)

new_cols = ncol(cepalstat_desempleo) - 1
cepalstat_desempleo = select(cepalstat_desempleo, 
                             1:new_cols)

new_cols = ncol(cepalstat_empleo) - 1
cepalstat_empleo = select(cepalstat_empleo, 
                          1:new_cols)

new_cols = ncol(cepalstat_remuneraciones) - 1
cepalstat_remuneraciones = select(cepalstat_remuneraciones, 
                                  1:new_cols)

new_cols = ncol(cepalstat_sector_financiero_monetario) - 1
cepalstat_sector_financiero_monetario = select(cepalstat_sector_financiero_monetario, 
                                               1:new_cols)

new_cols = ncol(cepalstat_sector_publico) - 1
cepalstat_sector_publico = select(cepalstat_sector_publico, 
                                  1:new_cols)

new_cols = ncol(cepalstat_BP_anual) - 1
cepalstat_BP_anual = select(cepalstat_BP_anual, 
                            1:new_cols)

new_cols = ncol(cepalstat_BP_trimestral) - 1
cepalstat_BP_trimestral = select(cepalstat_BP_trimestral, 
                                 1:new_cols)

new_cols = ncol(cepalstat_indicadores_derivados_de_la_BP) - 1
cepalstat_indicadores_derivados_de_la_BP = select(cepalstat_indicadores_derivados_de_la_BP, 
                                                  1:new_cols)

new_cols = ncol(cepalstat_deuda_externa) - 1
cepalstat_deuda_externa = select(cepalstat_deuda_externa, 
                                 1:new_cols)

new_cols = ncol(cepalstat_comercio_intrarregional) - 1
cepalstat_comercio_intrarregional = select(cepalstat_comercio_intrarregional, 
                                           1:new_cols)

new_cols = ncol(cepalstat_exp_imp_grandes_cat) - 1
cepalstat_exp_imp_grandes_cat = select(cepalstat_exp_imp_grandes_cat, 
                                       1:new_cols)

new_cols = ncol(cepalstat_exp_imp_pro_ppal_part_1_of_2) - 1
cepalstat_exp_imp_pro_ppal_part_1_of_2 = select(cepalstat_exp_imp_pro_ppal_part_1_of_2, 
                                                1:new_cols)

new_cols = ncol(cepalstat_exp_imp_pro_ppal_part_2_of_2) - 1
cepalstat_exp_imp_pro_ppal_part_2_of_2 = select(cepalstat_exp_imp_pro_ppal_part_2_of_2, 
                                                1:new_cols)

new_cols = ncol(cepalstat_exp_imp_servicios) - 1
cepalstat_exp_imp_servicios = select(cepalstat_exp_imp_servicios, 
                                     1:new_cols)

new_cols = ncol(cepalstat_exp_imp_totales_mensuales) - 1
cepalstat_exp_imp_totales_mensuales = select(cepalstat_exp_imp_totales_mensuales, 
                                             1:new_cols)

new_cols = ncol(cepalstat_exp_prim_manuf) - 1
cepalstat_exp_prim_manuf = select(cepalstat_exp_prim_manuf, 
                                  1:new_cols)

new_cols = ncol(cepalstat_indic_vol_precios_imp_exp) - 1
cepalstat_indic_vol_precios_imp_exp = select(cepalstat_indic_vol_precios_imp_exp, 
                                             1:new_cols)

new_cols = ncol(cepalstat_tipo_de_cambio) - 1
cepalstat_tipo_de_cambio = select(cepalstat_tipo_de_cambio, 
                                  1:new_cols)

new_cols = ncol(cepalstat_ipc_ipm_mensual) - 1
cepalstat_ipc_ipm_mensual = select(cepalstat_ipc_ipm_mensual, 
                                   1:new_cols)

new_cols = ncol(cepalstat_ipc_ipm_anual) - 1
cepalstat_ipc_ipm_anual = select(cepalstat_ipc_ipm_anual, 
                                 1:new_cols)

new_cols = ncol(cepalstat_mineria_manuf) - 1
cepalstat_mineria_manuf = select(cepalstat_mineria_manuf, 
                                 1:new_cols)

new_cols = ncol(cepalstat_turismo) - 1
cepalstat_turismo = select(cepalstat_turismo, 
                           1:new_cols)

new_cols = ncol(cepalstat_agricultura) - 1
cepalstat_agricultura = select(cepalstat_agricultura, 
                               1:new_cols)

new_cols = ncol(cepalstat_precios_combustibles) - 1
cepalstat_precios_combustibles = select(cepalstat_precios_combustibles, 
                                        1:new_cols)

new_cols = ncol(cepalstat_sector_real_mn_trimestral) - 1
cepalstat_sector_real_mn_trimestral = select(
  cepalstat_sector_real_mn_trimestral, 1:new_cols)
```

Finally, save the files as R binary data files (.rda). I'll show the code, but disable its evaluation, just to be careful,
because they form the base for alll the wrangling function and hence of the entire chain of analysis

```{r saving_raw_data_as_r, eval=FALSE}
save(cepalstat_sector_real_dolares_completo, 
     file = "./produced_data/cepalstat_sector_real_dolares_anual")

save(cepalstat_desempleo , file = "./produced_data/cepalstat_desempleo")

save(cepalstat_empleo , file = "./produced_data/cepalstat_empleo")

save(cepalstat_remuneraciones , file = "./produced_data/cepalstat_remuneraciones")

save(cepalstat_sector_financiero_monetario , 
     file = "./produced_data/cepalstat_sector_financiero_monetario")

save(cepalstat_sector_publico , file = "./produced_data/cepalstat_sector_publico")

save(cepalstat_BP_anual , file = "./produced_data/cepalstat_BP_anual")

save(cepalstat_BP_trimestral, file = "./produced_data/cepalstat_BP_trimestral")

save(cepalstat_indicadores_derivados_de_la_BP, 
     file = "./produced_data/cepalstat_indicadores_derivados_de_la_BP")

save(cepalstat_deuda_externa, file = "./produced_data/cepalstat_deuda_externa")

save(cepalstat_comercio_intrarregional, 
     file = "./produced_data/cepalstat_comercio_intrarregional")

save(cepalstat_exp_imp_grandes_cat, file = "./produced_data/cepalstat_exp_imp_grandes_cat")

save(cepalstat_exp_imp_pro_ppal_part_1_of_2, 
     file = "./produced_data/cepalstat_exp_imp_pro_ppal_part_1_of_2")

save(cepalstat_exp_imp_pro_ppal_part_2_of_2, 
     file = "./produced_data/cepalstat_exp_imp_pro_ppal_part_2_of_2")

save(cepalstat_exp_imp_servicios, file = "./produced_data/cepalstat_exp_imp_servicios")

save(cepalstat_exp_imp_totales_mensuales, 
     file = "./produced_data/cepalstat_exp_imp_totales_mensuales")

save(cepalstat_exp_prim_manuf, file = "./produced_data/cepalstat_exp_prim_manuf")

save(cepalstat_indic_vol_precios_imp_exp, 
     file = "./produced_data/cepalstat_indic_vol_precios_imp_exp")

save(cepalstat_tipo_de_cambio, file = "./produced_data/cepalstat_tipo_de_cambio")

save(cepalstat_ipc_ipm_mensual, file = "./produced_data/cepalstat_ipc_ipm_mensual")

save(cepalstat_ipc_ipm_anual, file = "./produced_data/cepalstat_ipc_ipm_anual")

save(cepalstat_agricultura, file = "./produced_data/cepalstat_agricultura")

save(cepalstat_mineria_manuf, file = "./produced_data/cepalstat_mineria_manuf")

save(cepalstat_turismo, file = "./produced_data/cepalstat_turismo")

save(cepalstat_precios_combustibles, 
     file = "./produced_data/cepalstat_precios_combustibles")

save(cepalstat_sector_real_mn_trimestral, 
     file = "./produced_data/cepalstat_sector_real_mn_trimestral")

```


### Reading data from the JEDH 
Fortunately the World Bank exposes data from the Joint External Debt Hub, of which it is a member, so we 
can use the `wbstats` package to download all data where the source correponds to JEDH. The list of countries 
it's give in our 33 and 20-members lists. This is implemented in  `get_joint_hub_data.R` and it's reproduced below:

```{r get_jedh_from_wb, cache=TRUE}
library(tidyverse)
library(wbstats)

load("./produced_data/cepal_33_countries")
load("./produced_data/cepal_20_countries")


# sources = updated_cache[[2]]$source
# JEDH_indicators = filter(updated_cache[[2]], source=="(JEDH) Joint External Debt Hub")
# JEDH_indicators_codes = JEDH_indicators$indicatorID
# 
# # shorter, using the fact we know now to look for JEDH as parte of the name of the source
# # long description is not included

JEDH_vars = wbsearch(pattern = "JEDH", cache=wb_cachelist, fields = "source")

debt_data_cepal_33 <- wb(country = cepal_33_countries[["iso3c"]],
                          indicator = JEDH_vars[["indicatorID"]],
                          mrv = 80)

debt_data_cepal_20 = debt_data_cepal_33 %>% 
                      filter(iso2c %in% cepal_20_countries[["iso2c"]])

save(debt_data_cepal_33, file = "./produced_data/debt_data_JEDH_cepal_33")
save(debt_data_cepal_20, file = "./produced_data/debt_data_JEDH_cepal_20")

```

This is the list of variables we get from that call:

```{r list_jedh_vars}
list_of_vars <- unique(debt_data_cepal_33$indicator)
print(list_of_vars)
```

To see how many countries have information on each variable and for what time span,
you will have to wait to read the data munging Rmarkdown document


### Monetary and banking data (EDD, Alejandra Acevedo)
We now turn to read data provided in a excel workbook with several sheets. The location 
of the data inside each sheet is not simple enough to use `readxl`, thus we employ the `XLConnect` 
package that let us to specify the exact region of the sheet we wish to extract.

The variable we retrive fromt his excel file, are:

- Préstamos Bancarios (Bank loans)
- Cartera vencida (Non performing loans)
- Crédito interno (internal credit)
- Meta de inflación, incluyendo máximos y mínimos (inflation target levels and upper-lower target bounds)
- Tasa de política monetaria (monetary policy rate)

The resulting R dataframe are stored in `datos_mon_finan_alej_messy.rda` for further manipulation.


The work is done in `read_monetarios_financieros_alejandra.R` which I reproduce below:

```{r read_monetarios, cache=TRUE}
library(tidyverse)
library(XLConnect)


wb = loadWorkbook("./raw_data/Datos_monetarios_de_Alejandra.xlsx")

sheet_name = "Tasa_politica"
rstart <-  3
rend <-  373
cstart <-  1
cend <- 34

tpm <- readWorksheet(wb, sheet = sheet_name, startRow = rstart,
                     endRow = rend, startCol = cstart, endCol = cend)



sheet_name = "Meta de Inflación"
rstart <-  5
rend <-  9
cstart <-  1
cend <- 180
meta_inf <- readWorksheet(wb, sheet = sheet_name, startRow = rstart,
                          endRow = rend, startCol = cstart, endCol = cend, header = FALSE)


rstart <-  12
rend <-  16
meta_inf_limsup <- readWorksheet(wb, sheet = sheet_name, startRow = rstart,
                          endRow = rend, startCol = cstart, endCol = cend, header = FALSE)

rstart <-  19
rend <-  23
meta_inf_liminf <- readWorksheet(wb, sheet = sheet_name, startRow = rstart,
                                 endRow = rend, startCol = cstart, endCol = cend, header = FALSE)


sheet_name = "Cartera Vencida"
rstart <-  3
rend <-  280
cstart <-  1
cend <- 34
cartera_vencida <- readWorksheet(wb, sheet = sheet_name, startRow = rstart,
                                 endRow = rend, startCol = cstart, endCol = cend)


sheet_name = "Crédito interno"
country_names_mess_ci <- readWorksheet(wb, sheet = sheet_name,
                               startRow = 1, endRow = 1,
                               startCol = 2, endCol = 199, header = FALSE) 

rstart <-  2
rend <-  323
cstarts <- seq(from = 3, by = 6, length.out = 33)
cends <- cstarts + 5

dfs_ci <- list_along(cstarts)

# autofitRow = FALSE is necessary to ensure all blocks have the same number of rows
# irrespective of the date of hte last observation recorded. Equal number of rows is 
# necessary for bind_rows() later
for (i in seq_along(cstarts) ) {
  
  dfs_ci[[i]] <- readWorksheet(wb, sheet = sheet_name,
                                 startRow = rstart, endRow = rend,
                                 startCol = cstarts[[i]], endCol = cends[[i]],
                                 autofitRow = FALSE) 
}



sheet_name = "Préstamos bancarios"
country_names_mess_pb <- readWorksheet(wb, sheet = sheet_name,
                                    startRow = 4, endRow = 4,
                                    startCol = 2, endCol = 240, header = FALSE) 

rstart <-  5
rend <-  349
cstarts <- seq(from = 2, by = 7, length.out = 33)
cends <- cstarts + 6

dfs_pb <- list_along(cstarts)

for (i in seq_along(cstarts) ) {
  dfs_pb[[i]] <- readWorksheet(wb, sheet = sheet_name,
                               startRow = rstart, endRow = rend,
                               startCol = cstarts[[i]], endCol = cends[[i]],
                               autofitRow = FALSE) 
}


save(tpm, 
     meta_inf, meta_inf_liminf, meta_inf_limsup,
     cartera_vencida, 
     dfs_pb, country_names_mess_pb, 
     dfs_ci, country_names_mess_ci, 
     file="./produced_data/datos_mon_finan_alej_messy")

```



### Reading KLEMS project data (EDD, Claudio Aravena)

Data for Argentina, Brazil, Chile, Colombia and Mexico. To make data comparable, 
data is aggregated up to 9 sectors.

The structure of each country-workbook is a follows:

- the fisrt sheet is called Notes, an contains a glossary of the variables present i the workbook, each of them stored in a separate sheet
- 36 sheets of numerical data, corresponding to 36 variables defined in Notes.

The reading script is called `read_klems_claudio.R` and saves its output to `klems_ar_br_cl_co_mx.rda` for further maipulation

```{r read_klems, cache=TRUE}
library(readxl)
library(tidyverse)

# workaround bc of the unwanted message DEFINEDNAME: blah blah ... blame the libxml package
quiet_read <- purrr::quietly(readxl::read_excel)
quiet_sheets <- purrr::quietly(readxl::excel_sheets)


## read chile
path_chi = "./raw_data/chi_output_09II.xls"
sheets_chi = quiet_sheets(path_chi)$result

klems_dfs_chi_map <- path_chi %>% 
                      map2(sheets_chi, function(x, y)  quiet_read(x, sheet = y) ) %>% 
                      map("result") 
names(klems_dfs_chi_map) <- sheets_chi


## read argentina
path_arg = "./raw_data/ar_output_09II.xls"
sheets_arg = quiet_sheets(path_arg)$result

klems_dfs_arg_map <- path_arg %>% 
  map2(sheets_arg, function(x, y)  quiet_read(x, sheet = y) ) %>% 
  map("result") 
names(klems_dfs_arg_map) <- sheets_arg


## read colombia
path_col = "./raw_data/co_output_09II.xls"
sheets_col = quiet_sheets(path_col)$result

klems_dfs_col_map <- path_col %>% 
  map2(sheets_col, function(x, y)  quiet_read(x, sheet = y) ) %>% 
  map("result") 
names(klems_dfs_col_map) <- sheets_col


## read brazil
path_bra = "./raw_data/br_output_09II.xls"
sheets_bra = quiet_sheets(path_bra)$result

klems_dfs_bra_map <- path_bra %>% 
  map2(sheets_bra, function(x, y)  quiet_read(x, sheet = y) ) %>% 
  map("result") 
names(klems_dfs_bra_map) <- sheets_bra


## read mexico
path_mex = "./raw_data/mex_output_09II_menor_agregacion.xls"
sheets_mex = quiet_sheets(path_mex)$result

klems_dfs_mex_map <- path_mex %>% 
  map2(sheets_mex, function(x, y)  quiet_read(x, sheet = y) ) %>% 
  map("result") 
names(klems_dfs_mex_map) <- sheets_mex

save(klems_dfs_arg_map, klems_dfs_bra_map, klems_dfs_chi_map, klems_dfs_col_map, 
     klems_dfs_mex_map, file="./produced_data/klems_ar_br_cl_co_mx")
```

### Reading Self-employment data (DDE, Jurgen Weller)

Very simple file cointaining  data on self employment as share of the total emplyment, 
for recent years, for 20 countries

```{r read_tcp}
library(tidyverse)
library(readxl)


TCP_2015 <- read_excel("./raw_data/TCP_2015.xlsx", skip = 1)
save(TCP_2015, file="./produced_data/trab_cuenta_propia_jurgen")

```


### Reading Balance of Payment information (DDE, Cecilia)
There may be substantial overlap with cepalstat data read before

```{r read_bop}
library(tidyverse)
library(stringr)
library(XLConnect)
library(lubridate)

bdpwb = loadWorkbook("./raw_data/BdpAnualSICValores10enero2017.xlsx")

## import tables from sheet = BP-America Latina
cstart <- 1
cend <-  43
cstarts <- rep(cstart, 43)
cends <- rep(cend, 43)

rstart <-  6
rend <-  61
row_block <-  rend - rstart
rstart_next <-  65
rows_to_next <-  rstart_next - rstart

number_of_tables <- 21

rstarts <-  seq(from = 6, by = rows_to_next, length.out = number_of_tables)
rends <- rstarts + row_block

dfs_bpal <- list_along(rstarts)

sheet_name <- "BP-América Latina"

for (i in seq_along(rstarts) ) {
  
  dfs_bpal[[i]] <- readWorksheet(bdpwb, sheet = sheet_name,
                              startRow = rstarts[[i]], endRow = rends[[i]],
                              startCol = cstarts[[i]], endCol = cends[[i]]) 
}



## import tables from sheet = BP-Caribe
cstart <- 1
cend <-  43
cstarts <- rep(cstart, 43)
cends <- rep(cend, 43)

rstart <-  6
rend <-  61
row_block <-  rend - rstart
rstart_next <-  65
rows_to_next <-  rstart_next - rstart

number_of_tables <- 14

rstarts <-  seq(from = 6, by = rows_to_next, length.out = number_of_tables)
rends <- rstarts + row_block

dfs_bpcar <- list_along(rstarts)

sheet_name <- "BP-Caribe"

for (i in seq_along(rstarts) ) {
  
  dfs_bpcar[[i]] <- readWorksheet(bdpwb, sheet = sheet_name,
                                 startRow = rstarts[[i]], endRow = rends[[i]],
                                 startCol = cstarts[[i]], endCol = cends[[i]]) 
}


save(dfs_bpal, dfs_bpcar, file = "./produced_data/from_BdP_excel")

```


### Reading export ad import data by origin/destinatio and by product (DITI, SC)

First by destination of exports and origin of imports (`read_expo_impo_origen_destino.R`):

```{r read_x_m_ori, cache=TRUE}
library(tidyverse)
library(stringr)
library(XLConnect)
library(lubridate)

x_m_oridest = loadWorkbook("./raw_data/expo_e_impo_por_destino_y_origen.xlsx")

num_of_countries = 16

all_dates = readWorksheet(x_m_oridest, sheet = 1, header = FALSE,
                          startRow = 2, endRow = 2, startCol = 2,
                          colTypes = "DATETIME")

all_dates_ch <- as.character.Date(all_dates, format = "%Y-%m-%d")

# all_dates <- cbind("dates", all_dates)


## read country names
rstart_names <- seq(from=3, by=21, length.out = num_of_countries)
all_names <- list_along(rstart_names)

for (i in seq_along(rstart_names) ) {
  
  this_name <- readWorksheet(x_m_oridest, sheet = 1, header = FALSE,
                             startRow = rstart_names[[i]],
                             endRow = rstart_names[[i]],
                             startCol = 1, endCol = 1) 
  
  all_names[[i]] <- this_name[[1]]
}


## read exports (x) data
rstart_x <- seq(from=5, by=21, length.out = num_of_countries)
n_x <- 7
rend_x <- rstart_x + n_x

dfs_x <- list_along(rstart_x)

for (i in seq_along(rstart_x) ) {
  
  this_df <- readWorksheet(x_m_oridest, sheet = 1, header = FALSE,
                          startRow = rstart_x[[i]], endRow = rend_x[[i]],
                          startCol = 1)
  
  this_df[5,1] <- "China"
  this_df[6,1] <- "Resto_de_Asia"
  
  colnames(this_df) <- append(c("region"), all_dates_ch)
  
  this_df$nombre_pais <- rep(all_names[[i]], n_x+1)
  
  this_df <- gather(this_df, date, value, -c(region, nombre_pais))
  
  dfs_x[[i]] <- this_df
}



## read imports (m) data
rstart_m <- seq(from=14, by=21, length.out = num_of_countries)
n_m <- 7
rend_m <- rstart_m + n_m

dfs_m <- list_along(rstart_m)

for (i in seq_along(rstart_m) ) {
  
  this_df <- readWorksheet(x_m_oridest, sheet = 1, header = FALSE,
                           startRow = rstart_m[[i]], endRow = rend_m[[i]],
                           startCol = 1)
  
  this_df[1,1] <- "Mundo"
  this_df[5,1] <- "China"
  this_df[6,1] <- "Resto_de_Asia"
  
  colnames(this_df) <- append(c("region"), all_dates_ch)
  
  this_df$nombre_pais <- rep(all_names[[i]], n_m+1)
  
  this_df <- gather(this_df, date, value, -c(region, nombre_pais))

  dfs_m[[i]] <- this_df
}

save(dfs_m, dfs_x, file = "./produced_data/x_m_por_origen_destino")

```


Now, by product groups (`read_expo_impo_ppales_productos.R`)

```{r read_x_m_prod, cache=TRUE}
library(tidyverse)
library(stringr)
library(XLConnect)
library(lubridate)

x_m_prod = loadWorkbook("./raw_data/expo_e_impo_por_grupo_de_productos.xlsx")


num_of_countries = 16

all_dates = readWorksheet(x_m_prod, sheet = 1, header = FALSE,
                          startRow = 3, endRow = 3, startCol = 2,
                          colTypes = "DATETIME")

all_dates_ch <- as.character.Date(all_dates, format = "%Y-%m-%d")

# all_dates <- cbind("dates", all_dates)


## read country names
rstart_names <- seq(from=4, by=12, length.out = num_of_countries)
all_names <- list_along(rstart_names)

for (i in seq_along(rstart_names) ) {
  
  this_name <- readWorksheet(x_m_prod, sheet = 1, header = FALSE,
                             startRow = rstart_names[[i]],
                             endRow = rstart_names[[i]],
                             startCol = 1, endCol = 1) 
  
  all_names[[i]] <- this_name[[1]]
}


## read exports (x) data
rstart_x <- seq(from=5, by=12, length.out = num_of_countries)
n_x <- 3
rend_x <- rstart_x + n_x

dfs_x_p <- list_along(rstart_x)

for (i in seq_along(rstart_x) ) {
  
  this_df <- readWorksheet(x_m_prod, sheet = 1, header = FALSE,
                          startRow = rstart_x[[i]], endRow = rend_x[[i]],
                          startCol = 1)
  
  # print(ncol(this_df))
  
  this_df[1,1] <- "total"
  
  colnames(this_df) <- append(c("producto"), all_dates_ch[1:ncol(this_df)-1])
  
  this_df$nombre_pais <- rep(all_names[[i]], n_x+1)
  
  this_df <- gather(this_df, date, value, -c(producto, nombre_pais))
  
  dfs_x_p[[i]] <- this_df
}



## read imports (m) data
rstart_m <- seq(from=10, by=12, length.out = num_of_countries)
n_m <- 4
rend_m <- rstart_m + n_m

dfs_m_p <- list_along(rstart_m)

for (i in seq_along(rstart_m) ) {
  
  this_df <- readWorksheet(x_m_prod, sheet = 1, header = FALSE,
                           startRow = rstart_m[[i]], endRow = rend_m[[i]],
                           startCol = 1)
  
  this_df[1,1] <- "total"
  
  colnames(this_df) <- append(c("producto"), all_dates_ch[1:ncol(this_df)-1])
  
  this_df$nombre_pais <- rep(all_names[[i]], n_m+1)
  
  this_df <- gather(this_df, date, value, -c(producto, nombre_pais))

  dfs_m_p[[i]] <- this_df
}

save(dfs_m_p, dfs_x_p, file = "./produced_data/x_m_por_producto")

```




### Read other BIS data (not JEDH)

There are three variables in this section:

- Debt service ratio for the private non-financial sector (in `dsr.xlsx`)

- Long series on total credit to the non-financial sectors (in `totcredit.xlsx`)

- Credit-to-GDP gap and underlying input data (in `c_gaps1612.xlsx`)


Contents of `read_bis_data.R`:

```{r read_bis, cache=TRUE}
library(tidyverse)
library(readxl)

c_gaps1612_qs <- read_excel("./raw_data/c_gaps1612.xlsx", 
                            sheet = "Quarterly Series", skip = 3)

c_gaps1612_documentation <- read_excel("./raw_data/c_gaps1612.xlsx", 
                                          sheet = "Documentation") 

c_gaps1612_content <- read_excel("./raw_data/c_gaps1612.xlsx", 
                             sheet = "Content")

totcredit_content <- read_excel("./raw_data/totcredit.xlsx", 
                                    sheet = "Content")

totcredit_documentation <- read_excel("./raw_data/totcredit.xlsx",
                    sheet = "Documentation")

totcredit_qs <- read_excel("./raw_data/totcredit.xlsx", 
                               sheet = "Quarterly Series", skip = 2)

dsr_content <- read_excel("./raw_data/dsr.xlsx", 
                      sheet = "Content")

dsr_documentation <- read_excel("./raw_data/dsr.xlsx", 
                                    sheet = "Documentation")

dsr_qs <- read_excel("./raw_data/dsr.xlsx", 
                       sheet = "Quarterly Series")


# after inspecting the data frames, get country names
names(c_gaps1612_documentation)
# backticks ` are necessary bc the variable name includes spaces 
c_gaps1612_countries <- unique(c_gaps1612_documentation$`Borrowers' country`)
# result: ARG, BRA, CHL and MEX

names(totcredit_documentation)
totcredit_countries <- unique(totcredit_documentation$`Borrowers' country`)
# result: ARG, BRA, CHL and MEX

names(dsr_documentation)
dsr_countries <- unique(dsr_documentation$`Borrowers' country`)
# result: BRA and MEX
```




